Reviewer Comments

### 2024-10-15 Manuscript.qmd
Further progress on review

{{< include Sections/01-introduction.qmd >}}
{{< include Sections/02-spatial-patterns.qmd >}}
{{< include Sections/03-connectivity.qmd >}}
{{< include Sections/04-sampling.qmd >}}
{{< include Sections/05-simulations.qmd >}}
{{< include Sections/06-genetics.qmd >}}
{{< include Sections/07-spatial-planning.qmd >}}
{{< include Sections/08-conclusion.qmd >}}

>>>>

{{< include Sections/01-introduction.qmd >}}
{{< include Sections/02-data.qmd >}}
{{< include Sections/03-spatial-patterns.qmd >}}
{{< include Sections/04-connectivity.qmd >}}
{{< include Sections/05-sampling.qmd >}}
{{< include Sections/06-simulations.qmd >}}
{{< include Sections/07-genetics.qmd >}}
{{< include Sections/08-spatial-planning.qmd >}}
{{< include Sections/09-conclusion.qmd >}}

### 2024-10-15 Review.qmd
Further progress on review

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

@@

>>>>

\textcolor[HTML]{0025ED}{Reply:}

@@

\textcolor[HTML]{0BB67D}{Reply: We agree that a general section on data could be useful for readers. Thus we added such a section at the beginning of the manuscript.}

@@

\textcolor[HTML]{0BB67D}{Reply: Indeed such a table would be very helpful. Thus, we removed all software-only paragraphs from the specific sections to avoid repetitions and added them to a table (Tab. 1).}

@@

\textcolor[HTML]{0BB67D}{Reply: We added a brief definition of edge length in this context.}

@@

\textcolor[HTML]{0BB67D}{Reply: We removed the term 'query' from the sentence because it potentially adds confusion. However, we believe that the sentence using areas instead is clear enough for readers to follow.}

@@

\textcolor[HTML]{0BB67D}{Reply: We agree that the whole section was confusing for readers unfamiliar with the topic. Thus, we reworded it to make it easier to follow also for all readers by removing a lot of the database jargon.}

@@

\textcolor[HTML]{0BB67D}{Reply: Thanks for pointing this out. We added it to the figure.}

@@

\textcolor[HTML]{0BB67D}{Reply: We added this correct specification that movement also includes plants. Furthermore, we believe that this aspect is also highlighted by the other suggestions of the reviewer that we implemented in this section.}

@@

\textcolor[HTML]{0BB67D}{Reply: We are unsure why the point about edge artifacts was added to the manuscript. Thus, we completely removed it here.}

@@

\textcolor[HTML]{0BB67D}{Reply: We agree that the original statement about underlying assumptions was unclear and reworded the sentence.}

@@

\textcolor[HTML]{0BB67D}{Reply: We added this specification.}

@@

\textcolor[HTML]{0BB67D}{Reply: We agree that this statement is true for most if not all methods presented in this manuscript. However, we believe it is also helpful to add here specifically for this section. Thus, we did not change it. However, we added a new short section at the beginning of the manuscript focusing on data specifically.}

@@

\textcolor[HTML]{0BB67D}{Reply: Thank you very much for adding this insightful remark about the references. We reworded the sentence to ensure they are reflected in the manuscript.}

@@

\textcolor[HTML]{0BB67D}{Reply: We added that digital twins benefit from the increasing rate of data generation in ecology as suggested.}

@@

\textcolor[HTML]{0BB67D}{Reply: Done as suggested.}

@@

\textcolor[HTML]{0BB67D}{Reply: We reworded the sentence being more specific about the references.}

@@

\textcolor[HTML]{0BB67D}{Reply: Thank you for pointing out this detail. We added this as another potentail drawback of the method.}

@@

\textcolor[HTML]{0BB67D}{Reply: We reworded the sentence stating that there are differences between solvers and that the optimal solver depends on the specific problem, however, could potentially not be accessible due to software licenses.}

### 2024-10-15 Sections/01-introduction.qmd
Further progress on review

<!-- R1: Very good idea to suggest software implementing methodological developments. It would be even better with a recapitulative table -->
Here, we introduce recent computational developments in some important topics of landscape ecology including software that implements or is potentially capable of implementing them (Fig. 1).
In this context, we are focusing on open-source software and specifically scripting languages such as R, Python, and Julia.
However, we do not provide a systematic literature review, or a general introduction to (computational) landscape ecology.
For a more comprehensive and general introduction to landscape ecology, please see @Turner2015LandscapeEcology, @Gergel2017LearningLandscape, or @With2019EssentialsLandscape.

>>>>

<!-- R1: Very good idea to suggest software implementing methodological developments. It would be even better with a recapitulative table -->
Here, we introduce recent computational developments in some important topics of landscape ecology (Fig. 1) including software that implements or is potentially capable of implementing them (Tab. 1).
In this context, we are focusing on open-source software and specifically scripting languages such as R, Python, and Julia.
However, we do not provide a systematic literature review, or a general introduction to (computational) landscape ecology.
For a more comprehensive and general introduction to landscape ecology, please see @Turner2015LandscapeEcology, @Gergel2017LearningLandscape, or @With2019EssentialsLandscape.

### 2024-10-15 Sections/03-spatial-patterns.qmd
Further progress on review

Software to calculate landscape metrics include the *landscapemetrics* [@Hesselbarth2019LandscapemetricsOpensource] and *multilandr* [@Huais2024MultilandrPackage] R packages, the *PyLandStats* Python library [@Bosch2019PyLandStatsOpensource], or the stand-alone software *FRAGSTATS* [@McGarigal2023FRAGSTATSV4].

@@

Based on concepts from thermodynamics initiated by the work of @Cushman2015ThermodynamicsLandscape and @Vranken2015ReviewUse, there has been a surge in the development of entropy-based metrics for landscape ecology in the last few years.
In the late 1800s, Boltzmann formulated a probabilistic interpretation of the second law of thermodynamics using the concepts of macrostate (the general state of a system) and microstate [the configuration of the system elements\; @Gao2019ComputationBoltzmann].
<!-- R1: the edge length -> Could you specify a bit -->
@Cushman2016CalculatingConfigurational proposed to relate the edge length to the microstate of the landscape, and use the proportion of microstates to compute the relative Boltzmann entropy of a landscape mosaic.
This approach was later generalized for calculations based on the raster surface model and point patterns [@Cushman2021EntropyLandscape].
Subsequently, @Gao2017HierarchybasedSolution proposed to use the Boltzmann entropy to quantify the complexity of a landscape surface by transforming the input raster into a series of landscape surfaces with different levels of detail (microstate) and calculating the Boltzmann entropy based on the total number of microstates that are able to generate the observed macrostate.
Recently, @Zhang2020CalculatingWasserstein extended the original definition of Boltzmann entropy to incorporate information about the adjacency of the same categories in the landscape mosaic by using the number of contiguous patches of the same category.
However, the relations between the Shannon entropy and the Boltzmann entropy and their interpretations in landscape ecology are the subject of ongoing discussion [e.g., @Stepinski2022CurbYour; @Huckeba2023MultiscaleSpatial].

@@

Entropy metrics are available in various software such as the R packages *landscapemetrics*, *bespatial* [@Nowosad2024BespatialBoltzmann], *rasterdiv* [@Rocchini2021RasterdivInformation], *LandComp* [@Konrad2023AdaptingMultiscale], or the Python package *PyLandStats*.

@@

Surface metrics are based on the gradient surface model using gridded raster data and continuous values [@McGarigal2009SurfaceMetrics; @Cushman2010GradientParadigm]. 
These metrics were mostly adapted from microscopy and molecular physics [@McGarigal2009SurfaceMetrics; @Kedron2018SurfaceMetrics].
<!-- R1: I guess that such metrics are also very scale dependent -->
The gradient surface model can increase the resemblance of the data to the natural world because it allows for the inclusion of more heterogeneity within each grid cell [@Cushman2010GradientParadigm; @Kedron2018SurfaceMetrics].
Surface metrics are able to quantify various characteristics, such as roughness, skewness and kurtosis, total and relative amplitudes, curvatures of local peaks, or surface bearing area ratios [@Cushman2010GradientParadigm].
Many surface metrics have analogous landscape metrics [@Cushman2010GradientParadigm], however, using surface metrics allows for the exploration of different or additional patterns and potentially pattern-process links [@McGarigal2009SurfaceMetrics; @Borthwick2020AlternativeQuantifications].
Nevertheless, similar to landscape metrics, surface metrics are also scale-dependent because they are used to quantify landscape heterogeneity, which is itself scale-dependent [@Frazier2016SurfaceMetrics].

Yet, software to calculate these metrics are still rare and further research is needed into the specific pattern-process links and ecologically meaningful interpretations [@Kedron2018SurfaceMetrics; @Frazier2019EmergingTrajectories].
Nonetheless, the *geodiv* R package [@Smith2021GeodivPackage] provides various surface metrics, and related textures from grey-level co-occurrence matrices are provided by the *glcm* R package [@Zvoleff2020GlcmCalculate].
Another approach to quantify the gradient surface model is based on frequencies and local adjacencies of continuous input pixel values as implemented in the *GraySpatCon* software [@Riitters2023MappingLandscape].

@@

The *vectormetrics* R package [@Matuszek2024VectormetricsLandscape] is specifically designed for vector metrics.
However, all general software frameworks designed for spatial data handling which support vector data can theoretically be used to calculate basic vector metrics.

@@

Spatial patterns can be analyzed through a range of computational operations, e.g., comparing, searching, or grouping.
These operations are based on spatial signatures (multi-numerical representations of landscape pattern) and dissimilarity measures (functions that quantify differences between the signatures).
Spatial patterns can be compared between different areas, or for the same area at two different moments in time.

Comparing spatial patterns is often used to analyze landscape dynamics, e.g., to detect changes in landscape structure over time [@Duncan2023MonitoringSpatial].
Furthermore, pattern-based search is a one-to-many operation in which the spatial signature of a query area is compared to the signatures of all other areas.
<!-- R1: As defined here, it looks to me that this operation is not necessarily linked to a query area, but rather to a spatial signature which can be directly specified or extracted from a query area -->
This allows for the identification of areas with similar patterns compared to the query area, e.g., areas with similar environmental conditions [@Jasiewicz2014LandscapeSimilarity].
Finally, the grouping of spatial patterns is a many-to-many operation in which spatial signatures are calculated for all areas and clustered based on their similarity [@Nowosad2021PatternbasedIdentification].

R, Python, and Julia have a wide range of packages for spatial operations, including *sf* [@Pebesma2018SimpleFeatures], or *terra* [@Hijmans2024TerraSpatial] in R, *GeoPandas* [@Jordahl2020GeopandasPython], or *rasterio* [@Gillies2013RasterioGeospatial] in Python, and *GeoStats.jl* [@Hoffimann2018GeoStatsjlHighperformance], or *GeoInterface.jl* [@JuliaGeoandcontributors2024GeoInterfaceJulia] in Julia.
These packages together with additional statistical and machine learning libraries allow for a wide range of spatial operations.
Additionally, the *motif* R package [@Nowosad2021MotifOpensource] provides a set of functions for spatial pattern operations, such as comparing, searching, or grouping spatial patterns customized for the use in landscape ecology.

>>>>

<!-- 
Software to calculate landscape metrics include the *landscapemetrics* [@Hesselbarth2019LandscapemetricsOpensource] and *multilandr* [@Huais2024MultilandrPackage] R packages, the *PyLandStats* Python library [@Bosch2019PyLandStatsOpensource], or the stand-alone software *FRAGSTATS* [@McGarigal2023FRAGSTATSV4].
-->

@@

Based on concepts from thermodynamics initiated by the work of @Cushman2015ThermodynamicsLandscape and @Vranken2015ReviewUse, there has been a surge in the development of entropy-based metrics for landscape ecology in the last few years.
In the late 1800s, Boltzmann formulated a probabilistic interpretation of the second law of thermodynamics using the concepts of macrostate (the general state of a system) and microstate [the configuration of the system elements\; @Gao2019ComputationBoltzmann].
<!-- R1: the edge length -> Could you specify a bit -->
@Cushman2016CalculatingConfigurational proposed to relate the edge length (defined as the length of neighboring cells with different LULC classes) to the microstate of the landscape, and use the proportion of microstates to compute the relative Boltzmann entropy of a landscape mosaic.
This approach was later generalized for calculations based on the raster surface model and point patterns [@Cushman2021EntropyLandscape].
Subsequently, @Gao2017HierarchybasedSolution proposed to use the Boltzmann entropy to quantify the complexity of a landscape surface by transforming the input raster into a series of landscape surfaces with different levels of detail (microstate) and calculating the Boltzmann entropy based on the total number of microstates that are able to generate the observed macrostate.
Recently, @Zhang2020CalculatingWasserstein extended the original definition of Boltzmann entropy to incorporate information about the adjacency of the same categories in the landscape mosaic by using the number of contiguous patches of the same category.
However, the relations between the Shannon entropy and the Boltzmann entropy and their interpretations in landscape ecology are the subject of ongoing discussion [e.g., @Stepinski2022CurbYour; @Huckeba2023MultiscaleSpatial].

@@

<!-- 
Entropy metrics are available in various software such as the R packages *landscapemetrics*, *bespatial* [@Nowosad2024BespatialBoltzmann], *rasterdiv* [@Rocchini2021RasterdivInformation], *LandComp* [@Konrad2023AdaptingMultiscale], or the Python package *PyLandStats*.
-->

@@

Surface metrics are based on the gradient surface model using gridded raster data and continuous values [@McGarigal2009SurfaceMetrics; @Cushman2010GradientParadigm]. 
These metrics were mostly adapted from microscopy and molecular physics [@McGarigal2009SurfaceMetrics; @Kedron2018SurfaceMetrics].
<!-- R1: I guess that such metrics are also very scale dependent -->
The gradient surface model can increase the resemblance of the data to the natural world because it allows for the inclusion of more heterogeneity within each grid cell [@Cushman2010GradientParadigm; @Kedron2018SurfaceMetrics].
Surface metrics are able to quantify various characteristics, such as roughness, skewness and kurtosis, total and relative amplitudes, curvatures of local peaks, or surface bearing area ratios [@Cushman2010GradientParadigm].
Many surface metrics have analogous landscape metrics [@Cushman2010GradientParadigm], however, using surface metrics allows for the exploration of different or additional patterns and potentially pattern-process links [@McGarigal2009SurfaceMetrics; @Borthwick2020AlternativeQuantifications].
Nevertheless, similar to landscape metrics, surface metrics are also scale-dependent because they are used to quantify landscape heterogeneity, which is itself scale-dependent [@Frazier2016SurfaceMetrics].
Yet, software to calculate these metrics are still rare and further research is needed into the specific pattern-process links and ecologically meaningful interpretations [@Kedron2018SurfaceMetrics; @Frazier2019EmergingTrajectories].
Another approach to quantify the gradient surface model is based on frequencies and local adjacencies of continuous input pixel values [@Riitters2023MappingLandscape].

<!--
Nonetheless, the *geodiv* R package [@Smith2021GeodivPackage] provides various surface metrics, and related textures from grey-level co-occurrence matrices are provided by the *glcm* R package [@Zvoleff2020GlcmCalculate].
Another approach to quantify the gradient surface model is based on frequencies and local adjacencies of continuous input pixel values as implemented in the *GraySpatCon* software [@Riitters2023MappingLandscape].
-->

@@

<!--
The *vectormetrics* R package [@Matuszek2024VectormetricsLandscape] is specifically designed for vector metrics.
However, all general software frameworks designed for spatial data handling which support vector data can theoretically be used to calculate basic vector metrics.
-->

@@

Spatial patterns can be analyzed through a range of computational operations, e.g., comparing, searching, or grouping.
These operations are based on spatial signatures (multi-numerical representations of landscape pattern) and dissimilarity measures (functions that quantify differences between the signatures), and can be calculated for different areas or for the same area at two different moments in time.

Comparing spatial signatures is often used to analyze landscape dynamics, e.g., to detect changes in landscape structure over time [@Duncan2023MonitoringSpatial].
<!-- R1: As defined here, it looks to me that this operation is not necessarily linked to a query area, but rather to a spatial signature which can be directly specified or extracted from a query area -->
Furthermore, signature-based search can be used to compare the spatial signature of a focal area to the spatial signatures of other areas.
This allows for the identification of areas with similar signatures compared to the focal area, e.g., areas with similar environmental conditions [@Jasiewicz2014LandscapeSimilarity].
Additionally, it is possible to calculate spatial signatures for multiple areas and group them to clusters based on their similarity of signatures [@Nowosad2021PatternbasedIdentification].

<!--
R, Python, and Julia have a wide range of packages for spatial operations, including *sf* [@Pebesma2018SimpleFeatures], or *terra* [@Hijmans2024TerraSpatial] in R, *GeoPandas* [@Jordahl2020GeopandasPython], or *rasterio* [@Gillies2013RasterioGeospatial] in Python, and *GeoStats.jl* [@Hoffimann2018GeoStatsjlHighperformance], or *GeoInterface.jl* [@JuliaGeoandcontributors2024GeoInterfaceJulia] in Julia.
These packages together with additional statistical and machine learning libraries allow for a wide range of spatial operations.
Additionally, the *motif* R package [@Nowosad2021MotifOpensource] provides a set of functions for spatial pattern operations, such as comparing, searching, or grouping spatial patterns customized for the use in landscape ecology.
-->

### 2024-10-15 Sections/04-connectivity.qmd
Further progress on review

<!-- R1: Here connectivity seems to be described through the prism of animals, you should modify it slightly to include plants, which do not move, but disperse through different vectors, e.g. animals, gravity, wind, or water -->
<!-- R1: facilitate or impeded movement and flow -->
Landscape connectivity describes how landscape features facilitate or impede movement [@Taylor2006LandscapeConnectivity] and can be divided into structural and functional connectivity.
It underpins ecosystem functionality, maintains biodiversity and populations, and plays an important role in many conservation actions [@CorreaAyram2016HabitatConnectivity].
Structural connectivity describes the physical arrangement of landscape elements, focusing on spatial aspects such as continuity and adjacency.
It is solely a landscape feature and independent of species characteristics [@Tischendorf2000UsageMeasurement; @CorreaAyram2016HabitatConnectivity].
Functional connectivity, in contrast, integrates landscape structure with the perceptual, behavioral, and dispersal characteristics of species, making it both species- and landscape-specific [@Adriaensen2003ApplicationLeastcost; @Taylor2006LandscapeConnectivity].

Although multiple approaches that measure connectivity have been developed [@Simpkins2018AssessingPerformance; @Keeley2021ConnectivityMetrics], the technical and conceptual quantification of connectivity has proven challenging [@Calabrese2004ComparisonshoppersGuide; @Taylor2006LandscapeConnectivity].
<!-- R1: movement and dispersal data -->
<!-- R1: that individuals preferentially move [or disperse] through -->
Generally, movement data from individuals is required to infer the elements in the landscape that individuals preferentially move through [@Simpkins2018AssessingPerformance].
While improving technologies have made tracking animal movement at high tempo-spatial resolution available [@Kays2015TerrestrialAnimal; @Nathan2022BigdataApproaches], it is still logistically challenging to track high numbers of individuals.
Thus, in practice connectivity is primarily determined by indirect estimations.
@Calabrese2004ComparisonshoppersGuide distinguished three types of estimates: i) structural connectivity determined by the physical attributes of the landscape, ii) potential connectivity as a combination of physical landscape attributes and limited information about dispersal characteristics of species, and iii) actual connectivity related to observations of individuals moving through a landscape.

@@

Many modern connectivity estimation approaches rely on resistance surfaces to represent the landscape [@Zeller2012EstimatingLandscape].
<!-- R1: to traverse or disperse through that cell -->
A resistance surface is a grid-based representation of a landscape where each cell is assigned a value representing the difficulty (or cost) for an individual of a particular species to traverse that cell based on landscape features such as habitat type, topography, or barriers [@Etherington2016LeastcostModelling].
<!-- R1: Please explain what "edge artifacts" are. I could not find any explanation in the cited reference -->
Creating resistance surfaces involves obtaining landscape data for the area of interest along with a sufficiently large buffer area to avoid edge artifacts [@Dutta2022OverviewComputational] and quantifying cost values for each cell using movement data.
This can include expert opinion [widely used due to low effort, however, difficult to measure accuracy\; @Krueger2012RoleExpert], detection data (single point locations of unknown individuals), relocation data (multiple sequential locations of the same individual but at low frequency), pathway data (high-frequency relocation data allowing for movement track inference), or genetic data (samples used to calculate genetic distances between populations).

<!-- R!: Least cost path is just the result of a mathematically formulated problem, so I don't think that it assumes anything. Assumptions are rather in how such results are used in an ecological context -->
Least-cost paths assume that individuals have perfect knowledge of the landscape and will choose the path that minimizes resistance costs perfectly with no deviations [@Adriaensen2003ApplicationLeastcost].
By calculating paths between two points, it is possible to estimate connectivity between these points based on the accumulated cost along the path.
This method can also generate accumulated cost surfaces depicting the minimum cost from a single point to all other locations, helping to identify reachable areas within a threshold.
The creation of least-cost paths is a well established technique and there are highly optimized and efficient algorithms available for its calculation.

@@

Within the landscape, discrete habitat patches are modeled as nodes, and the potential movement between patches is modeled as edges.
<!-- R1: Edges can be weighted or unweighted. They also can be directed or undirected -->
The edges can be binary, i.e., representing whether a connection between two nodes either exists or does not exists, or edges can be weighted by cost values representing movement efforts from one patch to another [@Petsas2021LandscapeConnectivity].
Connectivity can be quantified through various graph-theoretical metrics describing different aspects, such as the importance of specific patches in maintaining connectivity.
Widely reported connectivity metrics include the probability of connectivity index [@Saura2007NewHabitat], habitat availability metrics quantifying potential and functional connectivity [@Saura2010CommonCurrency; @Pietsch2018ContributionConnectivity], or the integral index of connectivity [@Pascual-Hortal2006ComparisonDevelopment] which is based on a binary connection model [@Saura2010CommonCurrency].

@@

A large number of graph theory based software applications have been developed to estimate connectivity including *ConScape.jl* [Julia\; @vanMoorter2023AcceleratingAdvances], *Omniscape.jl* [Julia\;@Landau2021OmniscapejlSoftware], *lconnect* [R\; @Mestre2023LconnectPackage], *grainscape* [R\; @Chubaty2020ToolboxGrainscape], *Circuitscape* [@Anantharaman2019CircuitscapeJulia], *Conefor* [@Saura2009ConeforSensinode], *Graphab* [@Foltete2012SoftwareTool], among others.

>>>>

<!-- R1: Here connectivity seems to be described through the prism of animals, you should modify it slightly to include plants, which do not move, but disperse through different vectors, e.g. animals, gravity, wind, or water -->
<!-- R1: facilitate or impeded movement and flow -->
Landscape connectivity describes how landscape features facilitate or impede movement, flow, and dispersal of organisms, e.g., direct movement of animals, but dispersal of sessile plants by wind or water, and can be divided into structural and functional connectivity [@Taylor2006LandscapeConnectivity].
It underpins ecosystem functionality, maintains biodiversity and populations, and plays an important role in many conservation actions [@CorreaAyram2016HabitatConnectivity].
Structural connectivity describes the physical arrangement of landscape elements, focusing on spatial aspects such as continuity and adjacency.
It is solely a landscape feature and independent of species characteristics [@Tischendorf2000UsageMeasurement; @CorreaAyram2016HabitatConnectivity].
Functional connectivity, in contrast, integrates landscape structure with the perceptual, behavioral, and dispersal characteristics of species, making it both species- and landscape-specific [@Adriaensen2003ApplicationLeastcost; @Taylor2006LandscapeConnectivity].

Although multiple approaches that measure connectivity have been developed [@Simpkins2018AssessingPerformance; @Keeley2021ConnectivityMetrics], the technical and conceptual quantification of connectivity has proven challenging [@Calabrese2004ComparisonshoppersGuide; @Taylor2006LandscapeConnectivity].
<!-- R1: movement and dispersal data -->
<!-- R1: that individuals preferentially move [or disperse] through -->
Generally, movement and dispersal data from individuals is required to infer the elements in the landscape that individuals preferentially move or disperse through [@Simpkins2018AssessingPerformance].
While improving technologies have made tracking animal movement at high tempo-spatial resolution available [@Kays2015TerrestrialAnimal; @Nathan2022BigdataApproaches], it is still logistically challenging to track high numbers of individuals.
Thus, in practice connectivity is primarily determined by indirect estimations.
@Calabrese2004ComparisonshoppersGuide distinguished three types of estimates: i) structural connectivity determined by the physical attributes of the landscape, ii) potential connectivity as a combination of physical landscape attributes and limited information about dispersal characteristics of species, and iii) actual connectivity related to observations of individuals moving through a landscape.

@@

Many modern connectivity estimation approaches rely on resistance surfaces to represent the landscape [@Zeller2012EstimatingLandscape].
<!-- R1: to traverse or disperse through that cell -->
A resistance surface is a grid-based representation of a landscape where each cell is assigned a value representing the difficulty (or cost) for an individual of a particular species to traverse or disperse that cell based on landscape features such as habitat type, topography, or barriers [@Etherington2016LeastcostModelling].
<!-- R1: Please explain what "edge artifacts" are. I could not find any explanation in the cited reference -->
Creating resistance surfaces involves obtaining landscape data for the area of interest, quantifying cost values for each cell using movement and dispersal data, and finally analyzing the surfaces [@Dutta2022OverviewComputational].
This can include expert opinion [widely used due to low effort, however, difficult to measure accuracy\; @Krueger2012RoleExpert], detection data (single point locations of unknown individuals), relocation data (multiple sequential locations of the same individual but at low frequency), pathway data (high-frequency relocation data allowing for movement track inference), or genetic data (samples used to calculate genetic distances between populations).

<!-- R!: Least cost path is just the result of a mathematically formulated problem, so I don't think that it assumes anything. Assumptions are rather in how such results are used in an ecological context -->
Least-cost modelling identifies potential paths between two points that minimizes the related movement or dispersal costs based on the resistance surface [@Adriaensen2003ApplicationLeastcost].
By calculating paths between two points, it is possible to estimate connectivity between these points based on the accumulated cost along the path.
This method can also generate accumulated cost surfaces depicting the minimum cost from a single point to all other locations, helping to identify reachable areas within a threshold.
The creation of least-cost paths is a well established technique and there are highly optimized and efficient algorithms available for its calculation.

@@

Within the landscape, discrete habitat patches are modeled as nodes, and the potential movement and dispersal between patches is modeled as edges.
<!-- R1: Edges can be weighted or unweighted. They also can be directed or undirected -->
The edges can be binary representing whether a connection between two nodes either exists or does not exists, weighted by cost values representing movement efforts from one patch to another, or undirected or directed allowing movement and dispersal in both or only one direction [@Urban2001LandscapeConnectivity; @Petsas2021LandscapeConnectivity].
Connectivity can be quantified through various graph-theoretical metrics describing different aspects, such as the importance of specific patches in maintaining connectivity.
Widely reported connectivity metrics include the probability of connectivity index [@Saura2007NewHabitat], habitat availability metrics quantifying potential and functional connectivity [@Saura2010CommonCurrency; @Pietsch2018ContributionConnectivity], or the integral index of connectivity [@Pascual-Hortal2006ComparisonDevelopment] which is based on a binary connection model [@Saura2010CommonCurrency].

@@

<!--
A large number of graph theory based software applications have been developed to estimate connectivity including *ConScape.jl* [Julia\; @vanMoorter2023AcceleratingAdvances], *Omniscape.jl* [Julia\;@Landau2021OmniscapejlSoftware], *lconnect* [R\; @Mestre2023LconnectPackage], *grainscape* [R\; @Chubaty2020ToolboxGrainscape], *Circuitscape* [@Anantharaman2019CircuitscapeJulia], *Conefor* [@Saura2009ConeforSensinode], *Graphab* [@Foltete2012SoftwareTool], among others.
-->

### 2024-10-15 Sections/05-sampling.qmd
Further progress on review

Software to perform multi-scale analysis include the R package *multilandr*, or the Python libraries *ZonalMetrics* [@Adamczyk2017ZonalMetricsPython], and *PyLandStats*.

>>>>

<!--
Software to perform multi-scale analysis include the R package *multilandr*, or the Python libraries *ZonalMetrics* [@Adamczyk2017ZonalMetricsPython], and *PyLandStats*.
-->

### 2024-10-15 Sections/06-simulations.qmd
Further progress on review

Pattern-based approaches simulate landscapes without assuming any underlying abiotic or biotic processes [i.e., neutral landscape model\; @Gardner1987NeutralModels].
The earliest neutral models are based on percolation theory and assign LULC classes randomly to cells in the landscape [@Gardner1987NeutralModels], or hierarchical models that consider different spatial scales while assigning cell values [@ONeill1992HierarchicalNeutral].
Landscapes characterized by continuous values can be simulated by fractal models, such as Brownian motion [@Palmer1992CoexistenceSpecies].
Borrowing from computer graphics, more recent neutral landscape models make use of spectral synthesis [e.g., Perlin noise\; @Etherington2022PerlinNoise], or binary space partitioning [@Etherington2022BinarySpace].
Neutral landscape models are also able to simulate landscapes dominated by anthropogenic activities based on least-cost networks [@Etherington2024LeastcostNetwork].
<!-- R1: Note that there, although the methods in reference 146 effectively rely on target values optimization (i.e. the model aims to get as close as possible to target values), the methods in reference 147 are rather based on target value satisfaction (the model guarantees that target values are strictly satisfied, or guarantees that they cannot be satisfied) -->
In order to ensure realistic neutral landscapes, several approaches exist that use comparisons with real landscapes [@Inkoom2017DesigningNeutral], or are based on target values optimization [@vanStrien2016ImprovedNeutral; @Justeau-Allaire2022FragmentedLandscape].
In order to make multiple neutral landscape model algorithms accessible, several software libraries are available including Python [NLMpy\; @Etherington2015NLMpyPython], R [NLMR\; @Sciaini2018NLMRLandscapetools], Julia [NeutralLandscapes.jl\; @Poisot2023NeutralLandscapesJulia], or stand-alone software [GradientLand\; @Cambui2015GradientLandSoftware].

@@

The digital twins concept include three connected elements, namely a physical object or system, its digital representation, and a data exchange between the former two [@Jones2020CharacterisingDigital; @deKoning2023DigitalTwins], and there is an increasing interest and use of digital twins in both the industry and academia [@Jones2020CharacterisingDigital].
The need for timely and evidence-based decision making in combination with increasing data availability make digital twins a powerful tool for landscape ecology [@deKoning2023DigitalTwins].
Digital twins are increasingly applied in agricultural landscapes research, e.g., for livestock farming, controlled environment farming, or fertilization management [@Cesco2023SmartAgriculture; @Purcell2023DigitalTwins].
Further applications include planning of rural ecological landscapes [@Tan2024DigitalTwin], or exploration of relationships between urban expansion and vegetation coverage [@Zhao2022ApplyingDigital].
<!-- R1: You could also mention the increasing availability and accessibility of sensors deployed by drones along with data processing tools (e.g. drone photogrammetry, LiDAR) which enable high-frequency, high-resolution monitoring of 2D and 3D landscape models -->
The use of general data processing, statistical and mechanistic models, or artificial intelligence allows digital twins to constantly update the digital representation and analyze, visualize, or predict the physical counterpart.
Thus, digital twins aim to provide real-time state of nature measurements, early detection of conservation trends, relationships of ecosystem trends and environmental conditions, impact assessments of interventions, or to identify uncertainties and information gaps [@deKoning2023DigitalTwins].

>>>>

Pattern-based approaches simulate landscapes without assuming any underlying abiotic or biotic processes [i.e., neutral landscape model\; @Gardner1987NeutralModels].
The earliest neutral models are based on percolation theory and assign LULC classes randomly to cells in the landscape [@Gardner1987NeutralModels], or hierarchical models that consider different spatial scales while assigning cell values [@ONeill1992HierarchicalNeutral].
Landscapes characterized by continuous values can be simulated by fractal models, such as Brownian motion [@Palmer1992CoexistenceSpecies].
Borrowing from computer graphics, more recent neutral landscape models make use of spectral synthesis [e.g., Perlin noise\; @Etherington2022PerlinNoise], or binary space partitioning [@Etherington2022BinarySpace].
Neutral landscape models are also able to simulate landscapes dominated by anthropogenic activities based on least-cost networks [@Etherington2024LeastcostNetwork].
<!-- R1: Note that there, although the methods in reference 146 effectively rely on target values optimization (i.e. the model aims to get as close as possible to target values), the methods in reference 147 are rather based on target value satisfaction (the model guarantees that target values are strictly satisfied, or guarantees that they cannot be satisfied) -->
In order to ensure realistic neutral landscapes, several approaches exist that use comparisons with real landscapes [@Inkoom2017DesigningNeutral], or are based on target values optimization [target values are optimized as close as possible\; @vanStrien2016ImprovedNeutral] or target value satisfaction [target values are strictly satisfied, or cannot be satisfied\; @Justeau-Allaire2022FragmentedLandscape].
<!-- In order to make multiple neutral landscape model algorithms accessible, several software libraries are available including Python [NLMpy\; @Etherington2015NLMpyPython], R [NLMR\; @Sciaini2018NLMRLandscapetools], Julia [NeutralLandscapes.jl\; @Poisot2023NeutralLandscapesJulia], or stand-alone software [GradientLand\; @Cambui2015GradientLandSoftware]. -->

@@

The digital twins concept include three connected elements, namely a physical object or system, its digital representation, and a data exchange between the former two [@Jones2020CharacterisingDigital; @deKoning2023DigitalTwins], and there is an increasing interest and use of digital twins in both the industry and academia [@Jones2020CharacterisingDigital].
The need for timely and evidence-based decision making in combination with increasing data availability make digital twins a powerful tool for landscape ecology [@deKoning2023DigitalTwins].
Digital twins are increasingly applied in agricultural landscapes research, e.g., for livestock farming, controlled environment farming, or fertilization management [@Cesco2023SmartAgriculture; @Purcell2023DigitalTwins].
Further applications include planning of rural ecological landscapes [@Tan2024DigitalTwin], or exploration of relationships between urban expansion and vegetation coverage [@Zhao2022ApplyingDigital].
<!-- R1: You could also mention the increasing availability and accessibility of sensors deployed by drones along with data processing tools (e.g. drone photogrammetry, LiDAR) which enable high-frequency, high-resolution monitoring of 2D and 3D landscape models -->
The use of general data processing, statistical and mechanistic models, or artificial intelligence allows digital twins to constantly update the digital representation and analyze, visualize, or predict the physical counterpart.
This is further facilitated by the increasing rate of data generation in ecology, e.g., due to sensors deployed on airplane, satellites, or unmanned aerial vehicle [@Farley2018SituatingEcology].
Thus, digital twins aim to provide real-time state of nature measurements, early detection of conservation trends, relationships of ecosystem trends and environmental conditions, impact assessments of interventions, or to identify uncertainties and information gaps [@deKoning2023DigitalTwins].
